---
title: "lab_03"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab that are in the data folder: the Maryland payments to vendors and Maryland grants and loans AND, using janitor, change the column names so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv", guess_max=369008)

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")

#clean
cleaned_payments <- payments |>clean_names()
cleaned_md_grants_loans <- md_grants_loans |>clean_names()

```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed in the payments data. Look at & search the records in the dataframe to start your search, starting with a broad search and then narrowing it down. Then write code to find rows where the vendor name contains some version of the county's school system.

HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1:

```{r}
pg_school <- cleaned_payments |>
  mutate(agency_name = str_to_upper(agency_name), vendor_name = str_to_upper(vendor_name)) |>
  filter(str_detect(vendor_name, "SCHOOL")) |>
  filter(str_detect(vendor_name, "PRINCE")|str_detect(vendor_name, "PG"))
pg_school
```


```{r}
pg_school |>
  summarise(
    paid_amount = sum(amount)
  )

```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2:

```{r}
pg_school |>
  group_by(fiscal_year)|>
  summarise(
    paid_amount = sum(amount)
  ) |>
  arrange(desc(paid_amount))

```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received.

Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe in words how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

The following 3 examples should be the same vendor. I would leave out "," and "." in the column name, remove every " INC" and "LLC" at the end of the vendor name
"A & I INC", "A&I INC", "A&I INC."
"AMERICAN DESIGN ASSOCIATE", "AMERICAN DESIGN ASSOCIATES", "AMERICAN DESIGN ASSOCIATES INC"
"CANNON DESIGN", "CANNON DESIGN INC"

A3.

```{r}
capitol_project_UMD <- cleaned_payments |>
  filter(str_detect(agency_name, "CAPITAL PROJECT"))|>
  filter(str_detect(agency_name, "MARYLAND"))

capitol_project_UMD |>
  group_by(vendor_name)|>
  summarise(
    paid_amount = sum(amount)
  ) |>
  arrange(vendor_name)
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which column(s) contain the information you want, your code should find just the grants for stem cell research in a single filter. Then calculate the total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.
A4: Amount of grant money to stem cell research projects peaks at 2009 and 2023, and hit the lowest during the 2021 pandemic.

```{r}
cleaned_md_grants_loans |>
  mutate(description = str_to_upper(description))|>
  filter(str_detect(description,"STEM")& str_detect(description,"CELL")) |>
  group_by(fiscal_year)|>
  summarise(grant_amount = sum(amount))

```

Q5: Using the Maryland grants and loan data, find records where the zip code is the one where the University of Maryland-College Park is located in and produce a list of grantees from that zip code, along with the total amount they received. Write a paragraph explaining how many grantees are there, what you would need to do to clean the results so that you had one record per grantee and which grantees are most interesting to you?
There're 14 grantees. There only 18 results after filtering so I might just clean it manually. I'll unify "UNIVERSITY OF MARYLAND", "UNIVERSITY OF MARYLAND COLLEGE PARK" and "UNIVERSITY OF MD,COLLEGE PARK" into "UNIVERSITY OF MARYLAND COLLEGE PARK." And make "UNIVERSITY OF MARYLAND - OFC. OF RESEARCH ADMIN. & ADVANCEMENT", "UNIV OF MD RESEARCH ADMINISTRATION & ADVANCEMENT" and "UNIVERSITY OF MARYLAND, DIVISION OF RESEARCH" into "UNIVERSITY OF MARYLAND - OFC. OF RESEARCH ADMIN. & ADVANCEMENT"
I'll look into the private companies that receive the highest amount of grant besides UMD. The highest one is Haystack Solutions, a cybersecurity company. The second and third ones, Zymetis, Inc and Remedium Technologies Inc., are both biotech companies.

A5:

```{r}
umdcp_grant <- cleaned_md_grants_loans |>
  filter(substr(zip_code, 1, 5) == 20742) 

umdcp_grant|>
  mutate(grantee = str_to_upper(grantee))|>
  group_by(grantee) |>
  summarise(grant_amount = sum(amount)) |>
  arrange(desc(grant_amount))

```
